Fast Algorithms for Intersection of Non-matching Grids Using \plucker Coordinates.

### 2016-12-12 paper_intersections.tex
Motivation improved.
Describe and discuss \epsilon.

The realistic models of groundwater processes including the transport processes and geomechanics have to deal with 
a complex nature of geological formations containing various small scale features as fractures (or fractured zones) and wells. 
Although of small scale, these features may have significant impact 
on the global behavior of the system and their representation in the numerical model is imperative. For example the fractures may form preferential 
paths which allow much faster transport that can not be fully captured by eqivalent continuum.
One possible approach is to model fractures and wells as lower dimensional objects and introduce their coupling with the surrounding continuum. 
The discretization then leads to the meshes of mixed dimensions, i.e. composed of elements of different dimension. This approach 
called mixed-dimensional analysis in the mechanics \cite{bournival_mesh-geometry_2008} is also studied in the groundwater context, see e.g. 
\cite{martin_modeling_2005}, \cite{fumagalli_numerical_2011}, \cite{brezina_analysis_2015} and 
already adopted by some groundwater simulation software, e.g FeFlow \cite{trefry_feflow:_2007} and Flow123d \cite{flow123d}.
Nevertheless as the complexity of the geometry increases (e.g. when lots of fractures are randomly generated) the compatible meshing becomes painful or even 
impossible. In order to avoid these difficulties we may discretize the continuum and every fracture and well independently, getting a non-matching 
(or incompatible) mesh of mixed dimensions and then apply XFEM to represent jumps of the solution on the fractures or singularities 
at the wells. The prerequisite for such approach is a fast and robust algorithm for calculating intersections of individual meshes.

@@

The calculation of the intersection proceeds as follows:
\begin{enumerate}
    \item Compute or reuse \plucker coordinates and permuted inner products: $\pi_p$, $\pi_{s_i}$, $\pi_p \odot \pi_{s_i}$, for $i=1,2,3$.
    \item Compute barycentric coordinates $w_i,\ i=1,2,3$ using \eqref{eq:bary_centric}.
    \item If any $w_i$ is less than $-\epsilon$ \notePE{I believe there was minus missing.}, there is no intersection, return empty IP. 
          Compare \fig{fig:plucker_products} and \eqref{eq:bary_centric} for better insight.
    \item If all $w_i$ are greater than $\epsilon$, we set $\tau_{S_2} = (2, 0)$ for the IC.
    \item If one $w_i$ is less than $\epsilon$, intersection is on the edge $s_i$, we set $\tau_{S_2} =(1,i)$.
    \item If two $w_i$ are less than $\epsilon$, intersection is at the vertex $V_i$, we set $\tau_{S_2}=(0,i)$.
    \item \label{item:coplanar} If all $w_i$ are less than $\epsilon$, the line is coplanar with the triangle, both objects are 
        projected to the plane $x_i=0$ where $i$ is the index of the maximal component of the triangle's normal vector.
        Every pair $p$, $s_i$ is checked for an intersection on $S_2$ boundary either inside $s_i$ or at a vertex $V_i$, setting the topological info $\tau_{S_2}$ to 
        $(1, i)$ or $(0, i)$, respectively. At most two ICs are obtained.
    \item For each IC the barycentric coordinates $(1-t, t)$ on the line $p$ are computed according to \eqref{eq:line}. 
    \item If $t\in (-\epsilon, \epsilon)$ or $t\in (1-\epsilon, 1+\epsilon)$,
        we set the end point of $S_1$: $\tau_{S_1} = (0,0)$ or $\tau_{S_1} = (0,1)$, respectively.
    \item If $t\notin (-\epsilon, 1+\epsilon)$, the IC is eliminated.
\end{enumerate}

We shall call every IC, that is created in the coplanar case, \emph{degenerate} since these will later need special treatment.

Checking the same sign of the permuted inner products can be viewed as a geometric predicate for the presence of the intersection and orientation of the 
line with respect to the triangle. Adaptive-precision evaluation of the geometric predicates was designed by Schewchuk \cite{shewchuk_adaptive_1997}
and used for 2d-2d mesh intersections in \cite{elsheikh_reliable_2012}. However, we rather apply a fixed tolerance check for the zero 
barycentric coordinates and consistently keep the topological positions in this and related algorithms.

>>>>

The realistic models of groundwater processes including the transport processes and geomechanics have to deal with 
a complex nature of geological formations containing various small scale features as fractures (or fractured zones) and wells. 
Although of small scale, these features may have significant impact 
on the global behavior of the system and their representation in the numerical model is imperative. For example the fractures may form preferential 
paths which allow much faster transport that can not be fully captured by eqivalent continuum.
One possible approach is to model fractures and wells as lower dimensional objects and introduce their coupling with the surrounding continuum. 
The discretization then leads to the meshes of mixed dimensions, i.e. composed of elements of different dimension. This approach 
called mixed-dimensional analysis in the mechanics \cite{bournival_mesh-geometry_2008} is also studied in the groundwater context, see e.g. 
\cite{martin_modeling_2005}, \cite{fumagalli_numerical_2011}, \cite{brezina_analysis_2015} and 
already adopted by some groundwater simulation software, e.g FeFlow \cite{trefry_feflow:_2007} and Flow123d \cite{flow123d}.
Nevertheless as the complexity of the geometry increases (e.g. when lots of fractures are randomly generated) the compatible meshing becomes painful or even 
impossible. In order to avoid these difficulties we may discretize the continuum and every fracture and well independently, getting a non-matching 
(or incompatible) mesh of mixed dimensions and then apply XFEM to represent jumps of the solution on the fractures or singularities 
at the wells. The prerequisite for such approach is a fast and robust algorithm for calculating intersections of individual meshes.
Although it is (currently) out of our interest the non-matching mesh approach allows a time evolving network of fractures necessary e.g. in 
modeling of hydraulic fracking.

@@

The calculation of the intersection proceeds as follows:
\begin{enumerate}
    \item Compute or reuse \plucker coordinates and permuted inner products: $\pi_p$, $\pi_{s_i}$, $\pi_p \odot \pi_{s_i}$, for $i=1,2,3$.
    \item Compute barycentric coordinates $w_i,\ i=1,2,3$ using \eqref{eq:bary_centric}.
    \item If any $w_i$ is less than $-\epsilon$ (see definition bellow), there is no intersection, return empty IP. 
          Compare \fig{fig:plucker_products} and \eqref{eq:bary_centric} for better insight.
    \item If all $w_i$ are greater than $\epsilon$, we set $\tau_{S_2} = (2, 0)$ for the IC.
    \item If one $w_i$ is less than $\epsilon$, intersection is on the edge $s_i$, we set $\tau_{S_2} =(1,i)$.
    \item If two $w_i$ are less than $\epsilon$, intersection is at the vertex $V_i$, we set $\tau_{S_2}=(0,i)$.
    \item \label{item:coplanar} If all $w_i$ are less than $\epsilon$, the line is coplanar with the triangle, both objects are 
        projected to the plane $x_i=0$ where $i$ is the index of the maximal component of the triangle's normal vector.
        Every pair $p$, $s_i$ is checked for an intersection on $S_2$ boundary either inside $s_i$ or at a vertex $V_i$, 
        setting the topological info $\tau_{S_2}$ to 
        $(1, i)$ or $(0, i)$, respectively. The ICs (at most two of them) obtained in this coplanar case will be called 
        \emph{degenerate} and will later need special treatment.
    \item For each IC the barycentric coordinates $(1-t, t)$ on the line $p$ are computed according to \eqref{eq:line}. 
    \item If $t\in (-\epsilon, \epsilon)$ or $t\in (1-\epsilon, 1+\epsilon)$,
        we set the end point of $S_1$: $\tau_{S_1} = (0,0)$ or $\tau_{S_1} = (0,1)$, respectively.
    \item If $t\notin (-\epsilon, 1+\epsilon)$, the IC is eliminated.
\end{enumerate}

The algorithm depends on the parameter $\epsilon$ used as a common tolerance parameter for detection of ICs with special positions. First, it is used 
in the sign check for permuted inner products, second, it is used for position check on the line. Only these two kinds of geometric predicates are used 
through the all intersection algorithms. Currently, we use just fixed value $\epsilon=10^{-9}$. It works on all realistic meshes we deal with, but we are aware 
of problems with possible unconsistent results of two different, but logically related predicates. Adaptive-precision evaluation of 
the geometric predicates was designed by Schewchuk \cite{shewchuk_adaptive_1997} and used for 2d-2d mesh intersections in \cite{elsheikh_reliable_2012}
in order to deal with these inconsistencies. It is a topic for future work to understand dependency between our geometric predicates
and decide if the adaptive-precision is the only way to guarantee the correctness of the algorithm even for various corner cases.
